Global Climate of the Ultra-Hot Jupiter WASP-103\lowercase{b} from \HST\ and Spitzer Phase Curve Observations

### 2017-10-09 ms.tex
models

We reduced the data from both programs using a custom pipeline developed for past analyses of WFC3 data \citep[for details see][]{kreidberg14a, kreidberg14b, kreidberg15a}. Briefly, we use the optimal extraction algorithm of \cite{horne86} to extract each up-the-ramp sample (or ``stripe") separately. The stripes are then summed to create the final spectrum. For each stripe, the extraction window is 24 pixels high and centered on the stripe midpoint. We estimate the background from the median of a region of the detector that is uncontaminated by the target spectrum (rows 5-50). The typical background counts are low (10-15 photoelectrons per pixel, roughly 0.03\% of the peak counts from the target star). We note that the extracted spectrum includes flux from a nearby star, which is separated from WASP-103 by less than two pixels \citep[0.2";][]{wollert15}. Our extracted spectrum includes flux from this star, which we account for later in the analysis.

@@

\section{Light Curve Fits}
We fit a two-component model to the full-orbit phase curves. One component models the planet signal (the thermal phase variation and transit), and the other component is a systematics model for changes in the instrument sensitivity. For each light curve, we fit the physical and systematic component simultaneously.

\subsection{Thermal Phase Variation Models}
We explored several different models for the planet's thermal phase variation. The simplest model was a sine curve, which has been used for many past analyses of exoplanet phase curves \citep[e.g.]{FIXME}.

We also fit climate maps using the open-source Python package \texttt{SPIDERMAN}. This package allows users to input a climate map (temperature or brightness as a function of latitude and longitude), and outputs a the thermal phase curve. The code is optimized to calculate

two-temperature model, with a uniform dayside temperature $T_d$ and a uniform nightside temperature $T_n$. also temperature map generated with spherical harmonics, the most general.  Physically-motivated kinematic model from \cite{zhang16}, which has just three free parameters (the nightside temperature $T_n$, the change in temperature from day-to-night side $\Delta_T$, and the ratio of radiative to advective timescales $\xi$). The kinematic model makes several simplifying assumptions (e.g, that the wind speed and direction are constant), but it reproduces GCM simulations fairly well.

We also fit a sine curve.

\subsubsection{Ellipsoidal Variations and Doppler Beaming}
ellipsoidal variations/doppler beaming

@@

\subsection{Systematics Models}
Both the \HST\ and \Spitzer\ phase curves have systematic noise caused by variations in the sensitivity of the instrument over time. For the WFC3 data, the dominant systematic is an orbit-long exponential trend due to charge traps filling up over successive exposures \citep{zhu17}. For \Spitzer\, the intrapixel sensitivity effect is the primary source of noise. The detector's pixels do not have uniform sensitivity, so slight changes in telescope pointing cause the recorded flux to vary.

@@

\noindent where $T(t)$ is a transit model, $E(t)$ is an eclipse model, $P$ is the planet orbital period, and $A$ and $\theta$ are free parameters to fit the amplitude and offset for the thermal phase variation. The free parameters for the transit and eclipse model are the planet-to-star radius ratio $r_p/r_s$, the planet-to-star flux ratio $f_p/f_s$, and a linear limb darkening parameter. The eclipse model is normalized such that the flux during eclipse is 1 and out of eclipse is $f_p/f_s$, and the transit model is normalized to unity out of transit. The planet's orbital parameters are poorly constrained by the WFC3 light curve due to incomplete phase coverage, so we fix the inclination to $i = 84.54^\circ$, the semi-major axis to stellar radius $a/R_s  = 2.925$, based on the best fit to the \emph{Spitzer} Channel 2 light curve (described below).  We use the orbital period $P = 0.925545613$ day, and time of inferior conjunction $t_0 = 2456836.2964455$ from \cite{southworth15}.  All free parameters were assumed to have a common value for both \emph{HST} visits, with the exception of $c$, $v_1$ and $v_2$, which were allowed to vary between visits. We explored adding an additional cosine term to model the thermal phase variation, but the additional degrees of freedom are not justified according to the Bayesian Information Criterion (BIC). The transit and eclipse models were calculated with the \texttt{batman} package \citep{kreidberg15a}. %Figure\;\ref{fig:w103_hst} shows the fitted broadband phase curve.

>>>>

We reduced the data from both programs using a custom pipeline developed for past analyses of WFC3 data \citep[for details see][]{kreidberg14a, kreidberg14b, kreidberg15b}. Briefly, we use the optimal extraction algorithm of \cite{horne86} to extract each up-the-ramp sample (or ``stripe") separately. The stripes are then summed to create the final spectrum. For each stripe, the extraction window is 24 pixels high and centered on the stripe midpoint. We estimate the background from the median of a region of the detector that is uncontaminated by the target spectrum (rows 5-50). The typical background counts are low (10-15 photoelectrons per pixel, roughly 0.03\% of the peak counts from the target star). We note that the extracted spectrum includes flux from a nearby star, which is separated from WASP-103 by less than two pixels \citep[0.2";][]{wollert15}. Our extracted spectrum includes flux from this star, which we account for later in the analysis.

@@

\section{Light Curve Fits}
We fit a two-component model to the full-orbit phase curves. One component models the astrophysical signal (the planet's thermal phase variation and transit), and the other component models the systematic noise introduced by changes in the instrument sensitivity. For each light curve, we fit the physical and systematic components simultaneously.

\subsection{Astrophysical Signal}
We assume the astrophysical signal $F_\mathrm{physical}(t)$ has the following form:

\begin{equation}
	F_\mathrm{physical}(t) = T(t) \times P(t) \times E(t)
\end{equation}

where $T(t)$ is the transit model, $P(t)$ is the thermal phase variation, and $E(t)$ is the ellipsoidal variation.

The transit model is calculated with the \texttt{batman} package \citep{kreidberg15a}. The free parameters for the transit fit are the planet radius $r_p$, orbital inclination $i$, the ratio of semi-major axis to stellar radius $a/R_s$, and a linear stellar limb darkening parameter $u$.  We fixed the orbital period $P = 0.925545613$ day and time of inferior conjunction $t_0 = 2456836.2964455$, based on \cite{southworth15}.  We found that the system parameters were poorly constrained for the \HST/WFC3 phase curve (due to incomplete phase coverage) and \Spitzer\ Channel 1 phase curve (due to correlated noise), so we fixed the the inclination to $i = 84.54^\circ$ and the semi-major axis to stellar radius $a/R_s  = 2.925$, based on the best fit to the \emph{Spitzer} Channel 2 light curve.

We explored several different models for the planet's thermal phase variation $P(t)$, using climate maps from the open-source Python package \texttt{SPIDERMAN}. This package allows users to input a climate map (temperature or brightness as a function of latitude and longitude), and generate the corresponding thermal phase curve over a specified wavelength range. We tested three different maps:  a two-temperature map, with a uniform dayside temperature $T_d$ and a uniform nightside temperature $T_n$; a map generated with spherical harmonics; and the physically-motivated kinematic model from \cite{zhang16}, which has just three free parameters (the nightside temperature $T_n$, the change in temperature from day-to-night side $\Delta_T$, and the ratio of radiative to advective timescales $\xi$). These models each have different strengths. The kinematic model incorporates the most prior knowledge of physics, and even though it makes several simplifying assumptions (e.g, that the wind speed and direction are constant), it reproduces GCM simulations fairly well.   The spherical harmonics map is the most flexible and makes the fewest assumptions about the true climate. The two-temperature model is the simplest and has the fewest free parameters.  
For comparison, we also fit the phase variation with a sinusoid, which has been used for many past analyses of exoplanet phase curves \citep[e.g.]{FIXME}.  We found FIXME.

In addition to the changing temperature of the planet's surface, the visible surface area also is expected to change as a function of orbital phase.  WASP-103b is so close to its host star that it is likely tidally distorted and partially filling its Roche lobe \citep{gillon14}, which introduces ellipsoidal variations $E(t)$ in the light curve.  Our light curves are not precise enough to detect the ellipsoidal variability, but we include a model prediction $E(t)$ to avoid introducing any bias in the thermal phase variation. We predict $E(t)$ using FIXME. FIXME doppler beaming?

@@

\subsection{Systematics}
Both the \HST\ and \Spitzer\ phase curves have systematic noise caused by variations in the sensitivity of the instrument over time. For the WFC3 data, the dominant systematic is an orbit-long exponential trend due to charge traps filling up over successive exposures \citep{zhu17}. For \Spitzer\, the intrapixel sensitivity effect is the primary source of noise. The detector's pixels do not have uniform sensitivity, so slight changes in telescope pointing cause the recorded flux to vary.

@@

\noindent where $T(t)$ is a transit model, $E(t)$ is an eclipse model, $P$ is the planet orbital period, and $A$ and $\theta$ are free parameters to fit the amplitude and offset for the thermal phase variation. The free parameters for the transit and eclipse model are the planet-to-star radius ratio $r_p/r_s$, the planet-to-star flux ratio $f_p/f_s$, and a linear limb darkening parameter. The eclipse model is normalized such that the flux during eclipse is 1 and out of eclipse is $f_p/f_s$, and the transit model is normalized to unity out of transit.

All free parameters were assumed to have a common value for both \emph{HST} visits, with the exception of $c$, $v_1$ and $v_2$, which were allowed to vary between visits. We explored adding an additional cosine term to model the thermal phase variation, but the additional degrees of freedom are not justified according to the Bayesian Information Criterion (BIC). The transit and eclipse models were calculated with the \texttt{batman} package \citep{kreidberg15a}. %Figure\;\ref{fig:w103_hst} shows the fitted broadband phase curve.
